---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---
#**Assignment**#

This assignment will be described in multiple parts. You will need to write a report that answers the questions detailed below. Ultimately, you will need to complete the entire assignment in a single R markdown document that can be processed by knitr and be transformed into an HTML file.

Throughout your report make sure you always include the code that you used to generate the output you present. When writing code chunks in the R markdown document, always use echo = TRUE so that someone else will be able to read the code. This assignment will be evaluated via peer assessment so it is essential that your peer evaluators be able to review the code for your analysis.

For the plotting aspects of this assignment, feel free to use any plotting system in R (i.e., base, lattice, ggplot2)

#**Setting up R Enviornment**#
```{r}
setwd("C:/Users/User/Desktop/Coursera/Reproducible Research/RepData_PeerAssessment1")
library(knitr)
opts_chunk$set(echo = TRUE, results = 'hold')
library(data.table)
library(ggplot2)
```
#**Loading and preprocessing the data**#

```{r}
#URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#download.file(URL, destfile = "/data/activity.zip")
unzip("data/activity.zip")
data <- read.csv("data/activity.csv", header = TRUE, sep = ",", colClasses = c("integer", "Date", "factor"))
data$month <- as.numeric(format(data$date, "%m"))
head(data)
dim(data)
```

#**What is mean total number of steps taken per day?**#

**Calculate the total number of steps taken per day**

```{r}
steps_day <- aggregate(steps ~ date, data, sum)
colnames(steps_day) <- c("Day","Steps")
```
**Make a histogram of the total number of steps taken each day**

```{r}
ggplot(data, aes(date, steps)) + geom_bar(stat = "identity", colour = "yellow", fill = "yellow", width = 0.7) + facet_grid(. ~ month, scales = "free") + labs(title = "Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
```
**Calculate and report the mean and median of the total number of steps taken per day**

```{r}
steps_mean   <- mean(steps_day$Steps, na.rm=TRUE)
steps_median <- median(steps_day$Steps, na.rm=TRUE)
```

The mean is `r steps_mean` and median is `r steps_median`

#**What is the average daily activity pattern?**#

**Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**

```{r}
steps_interval <- aggregate(data$steps, by = list(interval = as.numeric(as.character(data$interval))), FUN=mean, na.rm=TRUE)
names(steps_interval)[2] <- "StepsMean"

ggplot(steps_interval, aes(interval, StepsMean)) + geom_line(color = "yellow", size = 1) + labs(title = "Time Series Plot of the 5-minute Interval", x = "Interval", y = "Number of Steps")
```
**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**

max_steps <- steps_interval[steps_interval$StepsMean == max(steps_interval$StepsMean), ]

#**Imputing missing values**#

**Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)**

```{r}
missing_data <- sum(is.na(data$steps))
```

There are `r missing_data` missing values 

**Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.**

```{r}
updated_Data <- data 
for (i in 1:nrow(updated_Data)) {
    if (is.na(updated_Data$steps[i])) {
        updated_Data$steps[i] <- steps_interval[which(updated_Data$interval[i] == steps_interval$interval), ]$StepsMean
    }
}


updated_missing_data <- sum(is.na(updated_Data))

```
There are `r updated_missing_data` missing values 

**Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?**

```{r}
ggplot(updated_Data, aes(date, steps)) + geom_bar(stat = "identity", colour = "yellow", fill = "yellow", width = 1) + facet_grid(. ~ month, scales = "free") + labs(title = "Number of Steps per Day (with completed data)", x = "Date", y = "Total steps")

updated_steps_mean   <- mean(updated_Data$steps)
updated_steps_median <- median(updated_Data$steps)

```
**What is the impact of imputing missing data on the estimates of the total daily number of steps?**

The original mean is `r steps_mean` and median is `r steps_median`.
The updated (after filling in missing values) mean is `r updated_steps_mean` and median is `r updated_steps_median`.

#**Are there differences in activity patterns between weekdays and weekends?**#

**Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day**

```{r}
updated_Data$weekdays <- factor(format(updated_Data$date, "%A"))
levels(updated_Data$weekdays) <- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),weekend = c("Saturday", "Sunday"))
```

**Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).**

```{r}
StepsMean <- aggregate(updated_Data$steps, list(interval = as.numeric(as.character(updated_Data$interval)), weekdays = updated_Data$weekdays), FUN = "mean")
names(StepsMean)[3] <- "StepsMean"
library(lattice)
xyplot(StepsMean$StepsMean ~ StepsMean$interval | StepsMean$weekdays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Average Number of Steps")

```